---
title: 'colrow: Handling SimU, CR, and LU data'
author: "Pedro R. Andrade"
date: "`r format('21 November 2018')`"
urlcolor: blue
output:
  pdf_document:
    fig_caption: yes
    latex_engine: pdflatex
  html_document:
    df_print: paged
abstract: This package has functions to handle Simulation Units (SimU), ColRow (CR), and Large Unit (LU) data. It shows examples on how to export them as shapefiles, create maps, as well as summarize and plot them over time.
vignette: |
  %\VignetteEngine{knitr::rmarkdown} %\VignetteIndexEntry{colrow: Handling SimU, CR, and LU datas} %\VignetteEncoding{UTF-8}
---

# Introduction 

This package allows users to handle CR, SimU, and LU data. In order to
start, it is necessary to install the package from GitHub.

```{r, eval = FALSE}
devtools::install_github("pedro-andrade-inpe/colrow")
```

If the package will be installed in a Linux machine, it is necessary to
manually install some dependencies. Please visit the [webpage of the
package](http://www.github.com/pedro-andrade-inpe/colrow) in GitHub for more details.

# Creating shapefiles for a given country

First load the package.

```{r}
require(colrow)
```

Download world data from Dropbox
[here](https://www.dropbox.com/sh/sqocqe45jwmug2p/AAAbv-IAg24a_R4vYsP9zqV_a?dl=0)
and save them in a single directory. Create a variable in R to store this directory.

```{r}
dataDir <- "c:/Users/pedro/Dropbox/colrow/"
```

The content of the directory should have 17 files as shown below.

```{r}
list.files(dataDir)
```

See the available countries in alphabetical order, use `getCountries()`.
The code below shows the first 10.

```{r}
colrow::getCountries(dataDir)[1:10]
```

Then use `getLU()`, `getCR()`, and/or `getSimU()` to generate geospatial data for a given country.
They get as first argument the country name written in the same way as one of the outputs of `getCountries()`. The second argument is the directory where the data downloaded from Dropbox
was stored. These functions take some time to execute.

```{r}
country <- "Brazil"

myLU   <- colrow::getLU(country, dataDir)
myCR   <- colrow::getCR(country, dataDir)
mySimU <- colrow::getSimU(country, dataDir)
```

To see that everything is correct, plot the data.

```{r}
par(mfrow = c(1, 3), mar = c(5, 0.1, 5, 0.1))

plot(sf::st_geometry(mySimU), main = "SimU", col = "red"); box()
plot(sf::st_geometry(myCR),   main = "CR", col = "blue"); box()
plot(sf::st_geometry(myLU),   main = "LU", col ="green"); box()
```

Finally, save the data to be used by the processing functions. Each of this files
has a column named `ID` that will be used to match the objects from the CSV data.
Once these files are saved, it is not necessary to create them again every time
outputs are processed.

```{r}
sf::write_sf(myLU, paste0(country, "LU.shp"))
sf::write_sf(myCR, paste0(country, "CR.shp"))
sf::write_sf(mySimU, paste0(country, "SimU.shp"))
```

# Processing outputs

To process a CSV file, it is necessary to choose the created shapefile that matches
the representation used in the csv. Then, it is necessary to define the attribute names,
as the CSV files do not store a header.
These names must have one attribute called `ID` and another called `VALUE`.
Function `processFile()` gets the shapefile,
the CSV file, and the description of the attributes as arguments.
You can use `colrow::attrs()` to avoid quotes.

```{r}
csvfile <- system.file("extdata/scenarios/FC/Land_Compare3_FC.csv", package = "colrow")
attributes <- colrow::attrs(COUNTRY, ID, ALTI, SLP, SOIL, USE, SCENARIO, YEAR, VALUE)

result <- colrow::processFile(
  "BrazilCR.shp",
  csvfile,
  attributes
)
```

This function automatically ignores all attributes that have only one value (`COUNTRY`, `ALTI`,
`SLP`, `SOIL` and `SCENARIO` above). The other
attributes (except `ID` and `VALUE`) compose the created attribute names (`USE` and `YEAR` above). They are
concatenated according to their order in the CSV file.

```{r}
names(result)
```

It is possible to plot the created attributes by simply calling `plot()`, selecting
a given attribute.

```{r}
plot(result["CrpLnd2000"])
```

It is also possible to use `processFile()` to save the output when processing a file by adding
a fourth argument with the output file name. In this case, this function does not return anything.

```{r}
colrow::processFile(
  "BrazilCR.shp",
  csvfile,
  attributes,
  "brazilLandCompare.shp"
)
```

Note that shapefiles have limits: 255 or less columns, ten or less characters in the attribute
names. When working with large attribute names, it is necessary to simplify the names. In the
case below, seven names will be replaced. For instance, `CrpLnd` will be replaced by `cr` in the attribute names.

```{r}
convert <- list(
  CrpLnd = "cr", PriFor = "pr",
  NatLnd = "nl", ForReg = "fr",
  GrsLnd = "gl", MngFor = "mf",
  PltFor = "pl" 
)
```

If you want to simplify the years to two characters, just run:

```{r}
for(year in paste(seq(2000, 2050, 10))) # from 2000, 2010, ..., 2050
  convert[[year]] = substr(year, 3, 4) # to 00, 10, ..., 50

unlist(convert)
```

Finally, these simplifications can be used as fifth argument to `processFile()`.

```{r}
colrow::processFile(
  "BrazilCR.shp",
  csvfile,
  attributes,
  "brazilOutput.shp",
  convert
)
```

Sometimes there are more than one `VALUE` per unique `ID`, for each unique combination of
attributes. In this case, it is necessary to transform all the values into one single result.
As default, `processFile()` sums all the values, but it is possible to use any other function
passed as argument, such as `mean()`. Note that `sum()` is used for absolute values and
`mean()` for averages. To use another function, use argument `aggregate`, passing a 
function as value, as shown in the example below.

```{r}
result <- colrow::processFile(
  "BrazilCR.shp",
  system.file("extdata/csv/YIELD_COMPARE2.CSV", package = "colrow"),
  colrow::attrs(ID, CROP, ScenYear, VALUE),
  aggregate = sum # default value, could be omitted
)
```

# Plotting outputs

The example of plotting in this section uses `tmap` package.

```{r}
require(tmap)
```

In this example, we will draw the Brazilian Biomes on top of the ColRows. The shapefile
with this data is stored in file `br_biomes.shp` available within `colrow` package. We will
use CR data from `brazilLandCompare.shp`, created previously in this tutorial.

```{r}
brazil <- sf::read_sf("brazilLandCompare.shp")
biomes <- system.file("extdata/shape", "br_biomes.shp", package = "colrow") %>% sf::read_sf()
```

Then select a palette from the functions availables in `RColorBrewer` using `brewer.pal()`
(see the description of all colors available [here](http://colorbrewer2.org)) and
choose the cuts to represent the intervals of each slice. In the code below, as we have
seven cuts, we need to define six colors.

```{r}
max(brazil$GrsLnd2010)

cuts <- c(0, 50, 100, 150, 200, 250, 305)
rdPu <- RColorBrewer::brewer.pal(6, "RdPu")
```

Finally, use `tm_shape()`, `tm_fill()`, and `tm_borders()` to draw the map.

```{r}
tm_shape(brazil) +
  tm_fill(col = "GrsLnd2010", palette = rdPu, breaks = cuts, title = "Grass 2010") +
  tm_shape(biomes) +
  tm_borders(lwd = 1, col = "black")
```

If you want to zoom in the map to a given region, it is possible to set a `bbox` to
the `tm_shape()`:

```{r}
amazCerradoBox <- c(xmin = -74.5, xmax = -41.4, ymin = -25, ymax = 5.5) %>%
  sf::st_bbox(crs = st_crs(4326))

tm_shape(brazil, bbox = amazCerradoBox) +
  tm_fill(col = "GrsLnd2010", palette = rdPu, title = "Grass 2010") +
  tm_shape(biomes) +
  tm_borders(lwd = 1, col = "black")
```

# Handling csv files

You can handle csv files directly to create charts or to compute values without
creating shapefiles. First read data as a tibble using `readCSV()`:

```{r}
data <- colrow::readCSV(
  system.file("extdata/scenarios/FC/Land_Compare3_FC.csv", package = "colrow"),
  colrow::attrs(COUNTRY, ID, ALTICLASS, SLPCLASS, SOILCLASS, USE, SCENARIO, YEAR, VALUE)
)
```

To get the unique values per class, just use `unique()`:

```{r}
unique(data$USE)
unique(data$YEAR)
```

To check if the sum of all land use areas are the same for each year use some `dplyr` functions:

```{r}
result <- data %>%
  dplyr::group_by(YEAR) %>%
  dplyr::summarise(VALUE = sum(VALUE))

result
```

Then let us now compute the total of each land use area per year and rename the
attributes in order to use them directly in the plot.

```{r}
result <- data %>%
  dplyr::group_by(YEAR, USE) %>%
  dplyr::summarise(VALUE = sum(VALUE)) %>%  # sum by year/use
  dplyr::mutate(Year = YEAR, Use = USE) %>% # rename variables
  dplyr::mutate(Use = dplyr::recode(Use,    # rename attributes
    CrpLnd = "Crop Land",
    ForReg = "Forest Regrowth",
    GrsLnd = "Grass Land",
    MngFor = "Managed Forest",
    NatLnd = "Natural Land",
    PltFor = "Planted Forest",
    PriFor = "Primary Forest"))

result
```

Finally, just use `ggplot` to draw the evolution of each land use area per year.

```{r}
require(ggplot2)

ggplot(result) +
  aes(x = Year, y = VALUE, colour = Use) +
  geom_line(lwd = 1.5) +
  theme_bw() +
  ylab("Total (ha)")
```

# Final remarks

If you have suggestions or want to report an error, please visit the GitHup page [here](https://github.com/pedro-andrade-inpe/colrow).
